---
title: "Waterbird abundance percent change"
output: word_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```




```{r}

library(tidyverse)
library(here)
library(birdnames)
library(flextable)
library(ftExtra)
library(officer)


source(here("code/analysis_utilities.R"))



custom_bird_list <- readRDS("C:/Users/scott.jennings/OneDrive - Audubon Canyon Ranch/Projects/my_R_general/birdnames_support/data/custom_bird_list")


```

```{r eval = FALSE}
#Best model percent change estimates.



percent_changes <- readRDS(here("data_files/percent_changes")) %>% 
  mutate(common.name = factor(common.name, levels = c("All species combined", "Diving benthivore", "Herbivore", "Piscivore", "Brant", "Canada Goose", "Gadwall", "American Wigeon", "Mallard", "Northern Pintail", "Green-winged Teal", "Scaup spp", "Surf Scoter", "Black Scoter", "Bufflehead", "Common Goldeneye", "Common Merganser", "Red-breasted Merganser", "Ruddy Duck", "Pied-billed Grebe", "Horned Grebe", "Red-necked Grebe", "Eared Grebe", "Western/Clark's Grebe", "American Coot", "Forster's Tern", "Red-throated Loon", "Pacific Loon", "Common Loon", "Brandt's Cormorant", "Pelagic Cormorant", "Double-crested Cormorant", "Brown Pelican"))) %>% 
  arrange(common.name, phase, study.year) %>% 
  filter(!common.name %in% c("Diving benthivore", "Herbivore", "Piscivore"), phase == "overall") 



estimates_wide <- percent_changes %>% 
  select(common.name, which.year, pred.out) %>%  
  mutate(which.year = paste(which.year, "estimate", sep = ".")) %>% 
  pivot_wider(id_cols = c("common.name"), values_from = pred.out, names_from = which.year)


iucn <- readRDS(here("data_files/all_best_preds_response"))%>% 
    mutate(common.name = translate_bird_names(alpha.code, "alpha.code", "common.name")) %>%
    is_iucn_threatened_trend() %>% 
  ungroup() %>% 
  distinct(alpha.code, decline.dur)



out_table <- estimates_wide %>% 
  full_join(percent_changes %>% filter(!is.na(percent.change)) %>% distinct(common.name, alpha.code, percent.change)) %>% 
  full_join(iucn) %>% 
  left_join(read.csv(here("data_files/Foraging guild table.csv")) %>%
              mutate(alpha.code = translate_bird_names(common.name, "common.name", "alpha.code")) %>%
              dplyr::select(alpha.code, guild)) %>% 
  mutate(zbreak1 = "",
         zbreak2 = "",
         common.name = as.character(common.name)) %>% 
  select(common.name, guild, zbreak1, start.estimate, end.estimate, zbreak2, percent.change, decline.dur)

```

```{r eval=FALSE}

#Table 1. Changes in estimated abundance for 29 waterbird species and all species combined at Tomales Bay, California, 1992-2022. Estimates with 95% Confidence Intervals (%95 CI) are from the best supported model (lowest AICc value) for all but two species. In the case of Black Scoter, the best model (year2) gave estimates for the early years that seem unrealistically high, whereas the second best model (year; delta aicc = 0.3) seemed to give more reasonable estimates relative to the raw data. For Green-winged Teal the best model (Year + Restoration) estimates seem to be driven by a single particularly large value in 2011; 2 competitive models (Year, Δ AICc = 1.043; and Year + MOCI + Restoration, Δ AICc = 1.649) had mostly uninformative parameters (only Year in Year + MOCI + Restoration informative), so we included estimates for the next best model, the Intercept only model, which had Δ AICc = 2.152. Also shown is the foraging guild each species was assigned to and the number of years a species' decline met IUCN criteria for listing as near threatened or higher (blank if criteria not met). * indicates species we expected abundance to be related to the Giacomini Restoration.

flextable(out_table) %>% 
  add_header_row(values = c("", "Estimated abundance (95% CI)", ""), colwidths = c(3, 2, 3)) %>% 
  set_header_labels(common.name = "Species",
                    guild = "Foraging\nguild",
                    zbreak1 = "",
                    start.estimate = "1992",
                    end.estimate = "2022",
                    zbreak2 = "",
                    percent.change = "% change",
                    decline.dur = "# years decline met\nIUCN criteria")%>% 
  align(j = 2:8, align = "center", part = "all") %>% 
  border_remove() %>% 
  border(i = 2, j = 3:4, border.top = fp_border(color = "black"), part = "header") %>% 
  border(i = 1, border.top = fp_border(color = "black"), part = "header") %>% 
  border(i = 2, border.bottom = fp_border(color = "black"), part = "header") %>%
  border(i = nrow(out_table), border.bottom = fp_border(color = "black"), part = "body") %>%
  autofit() %>% 
  fit_to_width(max_width = 7.5) %>%
  colformat_md()
  


```

\newpage  


Model averaged percent change estimates

```{r}

mod_avg_percent_changes <- readRDS(here("data_files/mod_avg_per_change")) %>% 
  mutate(common.name = case_when(common.name == "all" ~ "All species combined",
                                 common.name == "diving benthivore" ~ "Diving benthivore",
                                 common.name == "herbivore" ~ "Herbivore",
                                 common.name == "omnivore" ~ "Omnivore",
                                 common.name == "piscivore" ~ "Piscivore",
                                 TRUE ~ as.character(common.name)),
         common.name = factor(common.name, levels = c("All species combined", "Diving benthivore", "Herbivore", "Piscivore", "Omnivore", "Brant", "Canada Goose", "Gadwall", "American Wigeon", "Mallard", "Northern Pintail", "Green-winged Teal", "Scaup spp", "Surf Scoter", "Black Scoter", "Bufflehead", "Common Goldeneye", "Common Merganser", "Red-breasted Merganser", "Ruddy Duck", "Pied-billed Grebe", "Horned Grebe", "Red-necked Grebe", "Eared Grebe", "Western/Clark's Grebe", "American Coot", "Forster's Tern", "Red-throated Loon", "Pacific Loon", "Common Loon", "Brandt's Cormorant", "Pelagic Cormorant", "Double-crested Cormorant", "Brown Pelican"))) %>% 
  arrange(common.name, study.year) %>% 
  filter(!common.name %in% c("Omnivore", "Diving benthivore", "Herbivore", "Piscivore")) 


year_spp <- readRDS(here("data_files/all_competitive_coefs")) %>% 
  filter(!alpha.code %in% c("Omnivore", "Diving benthivore", "Herbivore", "Piscivore")) %>% 
  filter(grepl("study.year", variable)) %>% 
  mutate(mod.type = ifelse(Delta_AICc == 0, "Best", "Competitive")) %>% 
  distinct(alpha.code, mod.type) %>%
  group_by(alpha.code) %>% 
  summarise(mods = paste(mod.type, collapse = ", "))
  
intercept_spp <- readRDS(here("data_files/all_competitive_coefs")) %>% 
  filter(!alpha.code %in% c("Omnivore", "Diving benthivore", "Herbivore", "Piscivore")) %>% 
  filter(Modnames == "intercept") %>% 
  mutate(int.mod = ifelse(Delta_AICc == 0, "Best", "Competitive")) %>% 
  distinct(alpha.code, int.mod) 


mod_avg_estimates_wide <- mod_avg_percent_changes %>% 
  select(common.name, which.year, pred.out) %>%  
  mutate(which.year = paste(which.year, "estimate", sep = ".")) %>% 
  pivot_wider(id_cols = c("common.name"), values_from = pred.out, names_from = which.year)


iucn <- readRDS(here("data_files/all_best_preds_response"))%>% 
    mutate(common.name = translate_bird_names(alpha.code, "alpha.code", "common.name")) %>%
    is_iucn_threatened_trend() %>% 
  ungroup() %>% 
  distinct(alpha.code, decline.dur)



mod_avg_out_table <- mod_avg_estimates_wide %>% 
  full_join(mod_avg_percent_changes %>% filter(!is.na(percent.change)) %>% distinct(common.name, alpha.code, percent.change)) %>% 
  full_join(iucn) %>% 
  left_join(read.csv(here("data_files/Foraging guild table.csv")) %>%
              mutate(alpha.code = translate_bird_names(common.name, "common.name", "alpha.code")) %>%
              dplyr::select(alpha.code, guild)) %>% 
  left_join(year_spp) %>% 
  left_join(intercept_spp) %>% 
  mutate(zbreak1 = "",
         zbreak2 = "",
         common.name = as.character(common.name),
         decline.dur = ifelse(is.na(mods) | !is.na(int.mod), NA, decline.dur)) %>% 
  select(common.name, zbreak1, start.estimate, end.estimate, zbreak2, percent.change, mods, int.mod, decline.dur, decline.dur)
```

```{r}
year_spp %>% 
  #mutate(best.comp = ifelse(mods == "Competitive", "Competitive", "Best")) %>% 
  group_by(mods)  %>% 
  summarise(num.spp = n(),
            spp = paste(alpha.code, collapse = ", ")) %>% 
  flextable() %>% 
  set_header_labels(mods = "Year in Best\nor Competitive",
                    num.spp = "# of species",
                    spp = "Species") %>% 
  width(width = c(1.5, 1, 4)) %>% 
  fit_to_width(max_width = 7.5) %>%
  colformat_md()
  
```

<br>

```{r}

mod_avg_out_table %>% 
  filter(!is.na(mods) & !is.na(int.mod)) %>% 
  group_by(mods, int.mod) %>% 
  summarise(num.spp = n(),
            spp = paste(common.name, collapse = ", ")) %>% 
  flextable() %>% 
  set_header_labels(num.spp = "# of species",
                    spp = "Species") %>% 
  width(width = c(1, 1, 1, 4)) %>% 
  fit_to_width(max_width = 7.5) %>%
  colformat_md()


```

<br>

```{r}
mod_avg_out_table %>% 
  mutate(true.change = is.na(int.mod) & !is.na(mods),
         up.down = case_when(true.change == TRUE & percent.change < 0 ~ "down",
                             true.change == TRUE & percent.change > 0 ~ "up",
                             true.change == FALSE ~ NA)) %>% 
  group_by(up.down, true.change) %>% 
  summarise(num.spp = n(),
            spp = paste(common.name, collapse = ", ")) %>% 
  ungroup() %>% 
  mutate(per.spp = 100*(num.spp/30),
         per.spp = round(per.spp, 1)) %>% 
  arrange(true.change) %>% 
  select(true.change, up.down, num.spp, per.spp, spp) %>% 
  flextable()  %>%
  set_header_labels(up.down = "Trend direction",
                    true.change = "Good evidence for\nabundance change",
                    num.spp = "# of species",
                    spp = "Species",
                    per.spp = "% of species") %>% 
  width(width = c(1, 1, 1, 1, 3)) %>% 
  fit_to_width(max_width = 7.5) %>%
  colformat_md()
           


```
<br>

```{r}
mod_avg_out_table %>% 
  mutate(true.change = is.na(int.mod) & !is.na(mods),
         iucn = !is.na(decline.dur)) %>% 
  filter(true.change == TRUE)%>% 
  group_by(iucn) %>% 
  summarise(num.spp = n(),
            spp = paste(common.name, collapse = ", ")) %>% 
  ungroup() %>% 
  mutate(per.spp = 100*(num.spp/22),
         per.spp = round(per.spp, 1)) %>% 
  select(iucn, num.spp, per.spp, spp) %>% 
  flextable()  %>%
  set_header_labels(iucn = "Change met\nIUNC criteria",
                    num.spp = "# of species",
                    spp = "Species",
                    per.spp = "% of species") %>% 
  width(width = c(1, 1, 1, 3)) %>% 
  fit_to_width(max_width = 7.5) %>%
  colformat_md()
           

```


Table 1. Model averaged estimates with 95% Confidence Intervals (%95 CI) for overall change in abundance for 29 waterbird species and all species combined at Tomales Bay, California, from 1992 to 2022. Also shown is the foraging guild each species was assigned to and the number of years a species' decline met IUCN criteria for listing as near threatened or higher (blank if criteria not met). * indicates species we expected abundance to be related to the Giacomini Restoration.
```{r}
flextable(mod_avg_out_table) %>% 
  add_header_row(values = c("", "Estimated abundance (95% CI)", ""), colwidths = c(2, 2, 5)) %>% 
  set_header_labels(common.name = "Species",
                    zbreak1 = "",
                    start.estimate = "1992",
                    end.estimate = "2022",
                    zbreak2 = "",
                    percent.change = "% change",
                    mods = "Models with\nYear effects",
                    int.mod = "Intercept-only\nmodel is:",
                    decline.dur = "# years decline met\nIUCN criteria") %>% 
  align(j = 2:9, align = "center", part = "all") %>% 
  border_remove() %>% 
  border(i = 2, j = 3:4, border.top = fp_border(color = "black"), part = "header") %>% 
  border(i = 1, border.top = fp_border(color = "black"), part = "header") %>% 
  border(i = 2, border.bottom = fp_border(color = "black"), part = "header") %>%
  border(i = nrow(mod_avg_out_table), border.bottom = fp_border(color = "black"), part = "body") %>%
  autofit() %>% 
  fit_to_width(max_width = 7.5) %>%
  colformat_md()
  


```

\newpage  

Comparing estimates from the best models to model averaged estimates.
```{r eval = FALSE}



bind_rows(out_table %>% select(common.name, percent.change, start.estimate, end.estimate) %>% mutate(which.estimate = "Best model"), 
          mod_avg_out_table %>% select(common.name, percent.change, start.estimate, end.estimate) %>% mutate(which.estimate = "Model averaged")) %>% 
  arrange(common.name) %>% 
  select(common.name, which.estimate, percent.change, start.estimate, end.estimate) %>% 
  mutate(common.name = ifelse(which.estimate == "Model averaged", "", common.name)) %>% 
  flextable() %>% 
  set_header_labels(common.name = "Species",
                    which.estimate = "Estimates from",
                    percent.change = "% change",
                    start.estimate = "1992",
                    end.estimate = "2022") %>% 
  autofit() %>% 
  fit_to_width(max_width = 7.5) %>%
  colformat_md() %>%
  border(i = seq(2, 60, by = 2), border.bottom = fp_border(color = "black"), part = "body") %>% 
  align(j = 2:4, align = "center", part = "all")
          


```